TacticToe: Learning to Prove with Tactics

### 2018-08-14 review_response.tex

\documentclass[]{scrartcl}
\usepackage{csquotes}
\MakeOuterQuote{"}

%opening
\title{Response for JAR submission}
\subtitle{TacticToe: Learning to Prove with Tactics}
\author{\mbox{Thibault Gauthier} \and \mbox{Cezary Kaliszyk} \and \mbox{Josef
		Urban} \and \mbox{Ramana Kumar} \and \mbox{Michael Norrish}}

@@

\maketitle

We thank the reviewers for the valuable comments and questions. We have 
attempted to correct all the issues suggested in the reviews and we give 
further answers to the questions raised in this response letter.

\section*{Reviewer 1}

\begin{leftbar}
No data is provided to help other researchers reproduce this experiment.  I 
consider this is a flaw of the paper.  I think the paper should be improved in 
the direction of making the experiment reproducible by other researchers, as 
much as possible. The paper provides numerical evaluation of its results, 
mostly in the form of 
percentage of success under a given timeout.   I would also be interesting to 
know how much it takes to re-run the whole process of learning and then using 
for the complete database.
\end{leftbar}

@@

\section*{Reviewer 2}

\begin{leftbar}
The only reservation I have with the paper is that it is not clear to me 
whether and how the proof tool is able to generate two commonly-used tactics. 
Firstly,
generating witnesses to existential goals using "EXISTS_TAC tm". If these 
tactics are out of scope for the proof tool, then are all goals containing 
existential
quantifiers impossible for the tool to prove? Secondly, reducing a goal using 
an implication theorem using "MATCH_MP_TAC th". Often such implication theorems 
are
highly specific to a single theory or even theorem, so how can tactics like 
this be effectively learned? In my opinion the paper would be improved by a 
brief
discussion on how these tactics could be learned, either in the present or a 
future system.
\end{leftbar}

A paragraph about predicting arrguments for these commonly used tactics has 
been added. more later

\section*{Reviewer 3}

\begin{leftbar}
The TacticToe system seems to particularly shine on inductive proofs over
recursive data structures, as the authors themselves note. It seems
therefore that the main place it wins over standard automated provers is
that it can recognize when and how to perform proof by induction. This
suggests that it may be less effective in other domains (e.g. where one has
to synthesize non-obvious witness terms) but also suggests to me that the
authors should also be comparing with the extensive literature on the
automation of inductive proof, e.g. Boyer-Moore, Bundy and "Proof
Planning", "inductionless induction" in rewriting, and the concrete
realizations in HOL family theorem provers by Boulton and Papapanagiotou.

The authors might also say a bit more about how much of this is specific to
HOL4, and how much would change if one wanted to implement something
similar in another prover either with a similar (HOL Light, Isabelle/HOL)
or different (ACL2, Coq, PVS) logic. While the authors give quite an
explicit discussion of the method, I did not see a link to the actual code.
Is this available?
\end{leftbar}

More details for the re

\end{document}

### 2018-08-14 tactictoe_jv.tex
minor + review response selected issues

Our prediction algorithms are not always accurate.
Therefore, a selected tactic may
fail, proceed in the wrong direction or even loop.
For that reason, predictions need to be accompanied by a
proof search mechanism that allows for backtracking and
can choose which proof tree to extend next and in which direction.
Our search algorithm is a Monte Carlo tree search~\cite{montecarlo} algorithm
that relies on a prior evaluation function and a prior policy function. Those 
priors are learned through direct supervised learning for the policy and via 
the data accumulated during orthogonalization for the evaluation. The
policy and evaluation are estimated by the simple supervised k-NN algorithm.
A characteristic of the MCTS algorithm is that it offers a good trade-off
between exploitation and exploration. This means that the algorithm searches
deeper the more promising branches and leaves enough time for the exploration
of less likely alternatives.

@@

\subsection{Supervised learning guidance}
Our MCTS algorithm relies on the guidance of two supervised learning methods.
The prior policy estimates the probability of success of a tactic $t$ on a
goal $g$ and associate it with the potential children that would be created by
the
application of $t$ to $g$. The prior evaluation judges if a branch is fruitful
by analyzing tactic
outputs. Both priors influence the rates at which branches of the search tree
are explored.

@@

Thanks to the larger search time available in this experiment, the timeout for
each tactic can be increased from the 0.02 seconds used in \tactictoe's initial
experiments~\cite{tgckju-lpar17} to 0.05
seconds. Removing tactics with similar effect as performed by the
orthogonalization process is only beneficial when running \tactictoe for short
period of time. It seems that the allotted time allows a strategy
running without orthogonalization to catch up. Yet, side experiments show
that orthogonalization becomes relevant again when we add additional
tactics through abstraction.
It is best to predict a list of 16 theorems to instantiate arguments of type
theorem list in tactics. At any rate, argument prediction for abstracted
tactics is the technique that has the highest impact on the success of
\tactictoe. Integration of ATPs being a specialization of this technique
contributes significantly as well. Experiments involving \eprover, run the ATP
using a separate process asynchronously with a timeout of 5 seconds and 128 premises.
The fact that \metis outperforms \eprover in this setting is due to the fact
that \metis is run with a very short timeout, allowing to close different part
of search tree quickly. Because \metis is weaker than \tactictoe as an ATP, it
is given less predictions. But even if an essential lemma for the proof of a
goal is not predicted, a modification on the goal performed by a tactic can
change its prediction to include the necessary lemma.
This effect minimizes the drawback of relying on a small number of predictions.

@@

\paragraph{Embellishment}
Without embellishment, the returned tactic is barely readable as it contains
information to guarantees that each \sml subterm is interpreted in the way in
any context. Since we return the proof at a point where the \sml interpreter is
in a specific state, we can strip unnecessary information, such as
module prefixes. If possible, we group all local declarations in the
proof under a single \texttt{let} binding at the start of the proof. We
also replace extended terms by their quoted version.
All in all, if a prettified tactic $t_p$ has the same effect its original
$t_o$, we replace $t_p$ by $t_o$ in the proof.\\

@@

It is important to consider the strength of the language used for building 
proofs and proof procedures when designing a system such as \tactictoe. Indeed, 
such system will be re-using those procedures.
In \isabelle, the language {\textsf{Eisbach}\xspace}~\cite{eisbach} was created 
on top of the standard tactic language by giving access to constructors of the 
tactic language to the end-user. The language 
{\textsf{PSL}\xspace}~\cite{NagashimaK17psl} was also developed. It can combine 
regular tactics and tools such as \sledgehammer.
In \coq, the \ltac~\cite{ltac} meta language was designed to enrich the 
tactic combinators of the \coq language. It adds functionalities such as 
recursors and matching operators for terms and proof contexts.
In \holfour, the strength of \sml as a proof strategy 
language has been demonstrated by the implementation of complex proof 
procedures such as \metis.

>>>>

Our prediction algorithms are not always accurate.
Therefore, a selected tactic may
fail, proceed in the wrong direction or even loop.
For that reason, predictions need to be accompanied by a
proof search mechanism that allows for backtracking and
can choose which proof tree to extend next and in which direction.
Our search algorithm is a Monte Carlo tree search~\cite{montecarlo} algorithm
that relies on a prior evaluation function and a prior policy function. Those 
priors are learned through direct supervised learning for the policy and via 
the data accumulated during orthogonalization for the evaluation. The
prior policy and evaluation are estimated by the simple supervised k-NN 
algorithm. We call them priors because these function are known a priori before 
starting the proof search. In contrast, we define later in this section a 
current policy and a current evaluation which are continuously updated by the 
MCTS algorithm during proof search through a feedback mechanism. A 
characteristic of the MCTS 
algorithm is that it offers a good 
trade-off between exploitation and exploration. This means that the algorithm 
searches deeper the more promising branches and leaves enough time for the 
exploration of less likely alternatives.

@@

\subsection{Supervised learning guidance}
Our MCTS algorithm relies on the guidance of two supervised learning methods.
The prior policy estimates the probability of success of a tactic $t$ on a
goal $g$ and associates it with the potential children that would be created by
the
application of $t$ to $g$. The prior evaluation judges if a branch is fruitful
by analyzing tactic
outputs. Both priors influence the rates at which branches of the search tree
are explored.

@@

Thanks to the larger search time available in this experiment, the timeout for
each tactic can be increased from the 0.02 seconds used in \tactictoe's initial
experiments~\cite{tgckju-lpar17} to 0.05
seconds. Removing tactics with similar effect as performed by the
orthogonalization process is only beneficial when running \tactictoe for a short
period of time. It seems that the allotted time allows a strategy
running without orthogonalization to catch up. Yet, side experiments show
that orthogonalization becomes relevant again when we add additional
tactics through abstraction.
It is best to predict a list of 16 theorems to instantiate arguments of type
theorem list in tactics. At any rate, argument prediction for abstracted
tactics is the technique that has the highest impact on the success of
\tactictoe. Integration of ATPs being a specialization of this technique
contributes significantly as well. Experiments involving \eprover, run the ATP
using a separate process asynchronously with a timeout of 5 seconds and 128 premises.
The fact that \metis outperforms \eprover in this setting is due to the fact
that \metis is run with a very short timeout, allowing to close different part
of search tree quickly. Because \metis is weaker than \tactictoe as an ATP, it
is given less predictions. But even if an essential lemma for the proof of a
goal is not predicted, a modification on the goal performed by a tactic can
change its prediction to include the necessary lemma.
This effect minimizes the drawback of relying on a small number of predictions.

@@

\paragraph{Embellishment}
Without embellishment, the returned tactic is barely readable as it contains
information to guarantee that each \sml subterm is interpreted in the same way 
in any context. Since we return the proof at a point where the \sml interpreter 
is
in a specific state, we can strip unnecessary information, such as
module prefixes. If possible, we group all local declarations in the
proof under a single \texttt{let} binding at the start of the proof. We
also replace extended terms by their quoted version.
All in all, if a prettified tactic $t_p$ has the same effect its original
$t_o$, we replace $t_p$ by $t_o$ in the proof.\\

@@

It is important to consider the strength of the language used for building 
proofs and proof procedures when designing a system such as \tactictoe. Indeed, 
such system will be re-using those procedures.
In \isabelle, the language {\textsf{Eisbach}\xspace}~\cite{eisbach} was created 
on top of the standard tactic language by giving access to constructors of the 
tactic language to the end-user. The language 
{\textsf{PSL}\xspace}~\cite{NagashimaK17psl} was also developed. It can combine 
regular tactics and tools such as \sledgehammer.
In \coq, the \ltac~\cite{ltac} meta language was designed to enrich the 
tactic combinators of the \coq language. It adds functionalities such as 
recursors and matching operators for terms and proof contexts.
In \holfour, the strength of \sml as a proof strategy 
language has been demonstrated by the implementation of complex proof 
procedures such as \metis. 
% TODO: Add a sentence about try-hard
